---
title: "Lab Modeling Mushrooms"
author: "Mark Mitri"
date: 21 September 2022
output: html_notebook
---


The goal for this lab is to become more comfortable with typical R coding tasks for data science modeling.

Put your name at the top as the author of this R notebook.

Start by reading the mushroom data set from http://archive.ics.uci.edu/ml/machine-learning-databases/mushroom/agaricus-lepiota.data and storing it in a data frame.
```{r}
shrooms <- read.csv("http://archive.ics.uci.edu/ml/machine-learning-databases/mushroom/agaricus-lepiota.data", sep = ",", header = FALSE)
shrooms$V17 <- NULL
shrooms
```

Sample two-thirds of the data for training, and one-third for testing

```{r}
shroomSize <- dim(shrooms)[1]
shroomSample <- sample(1:shroomSize, size=round(shroomSize/3), replace=FALSE, prob=rep(1/shroomSize, shroomSize))
shroom.train <- shrooms[-shroomSample,]; shroom.test <- shrooms[shroomSample,]
#shroom.test
#dim(shroom.train)
#dim(shroom.test)
```

Build an rpart decision tree to predict edible mushrooms
```{r}
library(rpart)
shroomTree <- rpart(V1 ~ ., data = shroom.train, method = "class")
shroomPred <- predict(shroomTree, shroom.test, type = 'class')
```

Calculate and print accuracy on the test data
```{r}
shroomAcc <- mean(shroomPred == shroom.test$V1)
print(paste('Accuracy of rpart: ', shroomAcc))
```

Do the same for C5.0.  Which one has better performance on the test data?
```{r}
#install.packages("C50")
library(C50)
shroom50 <- C5.0(as.factor(V1) ~ ., data=shroom.train, rules = TRUE)
shroom50pred <- predict(shroom50, newdata = shroom.test, type = 'class')
shroomTab <- table(data.frame(shroom.test$V1, pred = shroom50pred)); shroomTab
sum(diag(shroomTab))/sum(shroomTab)
```

