---
title: "HW7"
author: "Mark Mitri"
date: "10/25/2022"
output: html_notebook
---

(A total of 10 points.)

Consider a old-fashioned mail-order (only) catalog company that traditionally sends large holiday catalogs to two million past and one million potential customers. This is an expensive process, as the printing and mailing costs \$3 for each catalog, and typically only 6% of existing customers and 2% of potential customers make a purchase. The average sale to an existing customer nets the company \$55 in profit (i.e., after all expenses), while the average sale to a new customer is smaller and nets the company $40 in profit. Delivery of a catalog without a sale will result, of course, in an expense.

a, 4 points) What is the expected overall profit for this company to continue with their traditional plan?

```{r}
pastCust = 2000000
potCust = 1000000
cost = 3
pastPurchRate = 0.06
potPurchRate = 0.02
avgPastProfit = 55
avgPotProfit = 40

# Expected overall profit = = pR(x) * vR + [1 – pR(x)] * vNR

# First Try - Naive Calculation
EV1 = (pastPurchRate)*(avgPastProfit) + (potPurchRate)*(avgPotProfit)
# Second Try - Subtracting the cost per customer from the average Profit
EV2 = (pastPurchRate)*(avgPastProfit-(pastCust*cost)) + (potPurchRate)*(avgPotProfit-(potCust*cost))
# Third Try - Multiplying the average profit with the number of customers and then 
# subtracting the cost per customer from the average Profit
# I think this is the most accurate?
EV3 = (pastPurchRate)*((avgPastProfit*pastCust)-(pastCust*cost)) + (potPurchRate)*((avgPotProfit*potCust)-(potCust*cost))
# Fourth Try - Subtracting the cost of the customer not making a purchase
EV4 = (pastPurchRate)*((avgPastProfit*pastCust)-(pastCust*cost)) + (potPurchRate)*((avgPotProfit*potCust)-(potCust*cost)) - (1-pastPurchRate)*((avgPastProfit*pastCust)-(pastCust*cost)) + (1-potPurchRate)*((avgPotProfit*potCust)-(potCust*cost))
# Fifth Try - Naive Calculation * the number of customers
EV5 = (pastPurchRate)*(avgPastProfit*pastCust) + (potPurchRate)*(avgPotProfit*potCust)
# Sixth Try - Looking directly at the example of gambling on slides 14
EV6 = ((pastPurchRate)*((avgPastProfit*pastCust)-(pastCust*cost)) + (potPurchRate)*((avgPotProfit*potCust)-(potCust*cost))) - ((1-pastPurchRate)*(pastCust*cost) + (1-potPurchRate)*(potCust*cost))
# Seventh Try - Looking directly at the example of online marketing on slides 15
EV7 = (pastPurchRate * avgPastProfit - (1-pastPurchRate)*cost) + (potPurchRate * avgPotProfit - (1-potPurchRate)*cost)
# Eighth Try - Looking directly at the example of online marketing on slides 15 multiplying by the number of past and potential customers
EV8 = pastCust*(pastPurchRate * avgPastProfit - (1-pastPurchRate)*cost) + potCust*(potPurchRate * avgPotProfit - (1-potPurchRate)*cost)

EV2 <- prettyNum(EV2,big.mark=",",scientific=FALSE)
EV3 <- prettyNum(EV3,big.mark=",",scientific=FALSE)
EV4 <- prettyNum(EV4,big.mark=",",scientific=FALSE)
EV5 <- prettyNum(EV5,big.mark=",",scientific=FALSE)
EV6Pretty <- prettyNum(EV6,big.mark=",",scientific=FALSE)
EV8 <- prettyNum(EV8,big.mark=",",scientific=FALSE)

expProf <- c(EV1, EV2, EV3, EV4, EV5, EV6Pretty, EV7, EV8)
rowname <- c("1st Try (looks like a rate)", "2nd Try","3rd Try","4th Try","5th Try","6th Try (The correct Rate)", "7th Try (looks like a rate)", "Eighth Try")
results <- data.frame(expProf)
rownames(results) <- rowname
results
```

The company's data analytics department trained a model to predict (Y/N) who will make a holiday purchase from this catalog resulting in the following confusion matrices:

For existing customers
	p	n
Y	4	4
N	2	90

and

For potential customers
	p	n
Y	2	12
N	0	86

b, 6 points) You are tasked to answer the question of whether the company would make more money this season by only sending catalogs to the prospects that are predicted to purchase. Explain your reasoning by clearly showing your work as well as your final answer.  

```{r}
# Existing Customers
Total = 100

P1 = 6
probP1 = P1/Total
tpRate1 = 4/P1
fnRate1 = 2/P1

N1 = 94
probN1 = N1/Total
fpRate1 = 4/N1
tnRate1 = 90/N1

exProfExistCust = probP1*(tpRate1*55 + fnRate1*0) + probN1*(tnRate1*0 + fpRate1*-3)
moneyExist <- exProfExistCust*pastCust

# Potential Customers
P2 = 2
probP2 = P2/Total
tpRate2 = 2/P2
fnRate2 = 0/P2

N2 = 98
probN2 = N2/Total
fpRate2 = 12/N2
tnRate2 = 86/N2

exProfPotCust = probP2*(tpRate2*40 + fnRate2*0) + probN2*(tnRate2*0 + fpRate2*-3)
moneyPot <- exProfPotCust*potCust

totalMoney = moneyExist + moneyPot

exProfExistCust <- prettyNum(exProfExistCust,big.mark=",",scientific=FALSE)
exProfPotCust <- prettyNum(exProfPotCust,big.mark=",",scientific=FALSE)
moneyExist <- prettyNum(moneyExist,big.mark=",",scientific=FALSE)
moneyPot <- prettyNum(moneyPot,big.mark=",",scientific=FALSE)
totalMoney <- prettyNum(totalMoney,big.mark=",",scientific=FALSE)

Results <- c(exProfExistCust, exProfPotCust, moneyExist, moneyPot, totalMoney)
rowname <- c("Expected Rate of Profit for Existing Customers", "Expected Rate of Profit for Potential Customers", "Expected Profit for Existing Customers", "Expected Profit for Potential Customers", "Total Profit for Existing Customers")
moneyResults <- data.frame(Results)
rownames(moneyResults) <- rowname
moneyResults
```


Submission:

- Make sure this notebook is cleanly written so that someone can simply run the whole thing.  Submit this complete notebook via coursesite.
- Make sure to comment sufficiently that a reader will understand what you are doing.
