---
title: "HW6"
author: "Mark Mitri"
date: "10/17/2022"
output: html_notebook
---

(A total of 10 points.)

Consider the problem of time series analysis for parking data.  At the webpage https://archive.ics.uci.edu/ml/datasets/Parking+Birmingham you'll find a dataset about parking utilization.  You'll need to download the zip file and unzip it locally.

```{r}
# YOUR CODE HERE
```

After you have read in the data, and examined it, you'll find that it contains occupancy numbers for multiple locations.  Let's work with just one location: 
SystemCodeNumber == "BHMBCCMKT01" so generate a new dataframe by extracting the subset for this location.  (2 points)

Let's build a model of regression, to predict the next occupancy level reading.  Using the second time series analysis R notebook as reference, create a dataframe with multiple attributes (3 points).  Note that this dataset is quite different, and so not all of the attributes we saw in class could be created here.

```{r}
# YOUR CODE HERE
```

Build a regression model based on those attributes (3 points) and evaluate it using the simple measures we saw in class (2 points).  Are you able to get 90% correlation?

```{r}
# YOUR CODE HERE
```

Submission:

- Make sure this notebook is cleanly written so that someone can simply run the whole thing.  Submit this complete notebook via coursesite.
- Make sure to comment sufficiently that a reader will understand what you are doing.
