---
title: "HW6"
author: "Mark Mitri"
date: "10/17/2022"
output: html_notebook
---

(A total of 10 points.)

Consider the problem of time series analysis for parking data.  At the webpage https://archive.ics.uci.edu/ml/datasets/Parking+Birmingham you'll find a dataset about parking utilization.  You'll need to download the zip file and unzip it locally.

```{r}
# Read in the local data as a csv with headers
parking <- read.csv("C:\\Users\\MitriDell\\Documents\\Textbooks\\Senior - Fall 2022\\CSE 160\\Homework\\HW6\\dataset.csv", header = TRUE)

# Visualize the data
head(parking)
```

After you have read in the data, and examined it, you'll find that it contains occupancy numbers for multiple locations.  Let's work with just one location: 
SystemCodeNumber == "BHMBCCMKT01" so generate a new dataframe by extracting the subset for this location.  (2 points)

Let's build a model of regression, to predict the next occupancy level reading.  Using the second time series analysis R notebook as reference, create a dataframe with multiple attributes (3 points).  Note that this dataset is quite different, and so not all of the attributes we saw in class could be created here.

```{r}
# Select only the rows of data the match the SystemCodeNumber of BHMBCCMKT01
parkSub <- parking[parking$SystemCodeNumber == "BHMBCCMKT01",]

# Create a separate dataframe for making our attributes
parkdf <- data.frame(parkSub=parkSub)

# Create an attribute to predict the future Occupancy based on the previous. Have to make the first element NA
parkdf$Prior1[1] <- NA

# Loop through all the rows and record the previous Occupancy number
for(i in 2:1312){
  parkdf$Prior1[i] <- parkdf$parkSub.Occupancy[i-1]
}

# Create an attribute to predict the future Occupancy based on the 3 previous. Have to make the first 3 elements NA
parkdf$Prior3[1] <- NA
parkdf$Prior3[2] <- NA
parkdf$Prior3[3] <- NA

# Loop through all the rows and record the previous 3 Occupancy number
for(i in 4:1312){
  parkdf$Prior3[i] <- parkdf$parkSub.Occupancy[i-2]
}

# Create an attribute to track the time of day the occupancy usually increases. A repeated pattern that loops 18 times from ~7 am - 5 pm
parkdf$Seasonal <- rep(1:18, each=1, length.out=1312)

# A number to roughly represent the time that the data was taken
parkdf$SimpleTime <- rep(7:16, each=2, length.out=1312)
```

Build a regression model based on those attributes (3 points) and evaluate it using the simple measures we saw in class (2 points).  Are you able to get 90% correlation?

```{r}
# Calculate the number of rows parkdf has
rows <- dim(parkdf)[1]
#rows

# Separate the data into 80% train & 20% test
train <- parkdf[13:floor(12+(rows-12)*.8),]
test <- parkdf[-(1:floor(12+(rows-12)*.8)),]

#dim(train)
#dim(test)

# Create the linear regression model using the training data on the attribute Occupancy based on the previous occupancy
model <- glm(parkSub.Occupancy ~ Prior1 + Prior3 + Seasonal + SimpleTime, data = train)
summary(model)
pred.value <- predict(model,test)

# Make a data frame to hold the test results
actuals_preds <- data.frame(cbind(actuals=test$parkSub.Occupancy, predicteds=pred.value))
correlation_accuracy <- cor(actuals_preds$actuals, actuals_preds$predicteds)

# Print the accuracy of the model
print("Correlation Accuracy is:")
correlation_accuracy
```

Submission:

- Make sure this notebook is cleanly written so that someone can simply run the whole thing.  Submit this complete notebook via coursesite.
- Make sure to comment sufficiently that a reader will understand what you are doing.
