---
title: "Lab Modeling Mushrooms"
author: Brian Davison
date: 17 July 2022
output: html_notebook
---


The goal for this lab is to become more comfortable with typical R coding tasks for data science modeling.

Start by reading the mushroom data set from http://archive.ics.uci.edu/ml/machine-learning-databases/mushroom/agaricus-lepiota.data 
```{r}
mushrooms <- read.csv("http://archive.ics.uci.edu/ml/machine-learning-databases/mushroom/agaricus-lepiota.data", header=F)
```

Sample two-thirds of the data for training, and one-third for testing

```{r}
mushrooms$V17 <- NULL # removed because it was constant, and some kinds of models complain

for (i in 1:ncol(mushrooms)) {
  mushrooms[[i]] <- factor(mushrooms[[i]])
}

length <- nrow(mushrooms)
set.seed(7)  # C5.0 gets 100% accuracy with a seed of 88
s <- sample(1:length, floor(length/3))
test <- mushrooms[s,]
train <- mushrooms[-s,]
```

Build an rpart decision tree to predict edible mushrooms
```{r}
library(rpart)
tree <- rpart(V1 ~ ., data=train, method="class")
```

Calculate and print accuracy on the test data
```{r}
pred <- predict(tree, test, type="class")
tab <- table(test$V1, pred)
tab
print(paste("Accuracy:", sum(diag(tab)) / sum(tab)))
```

Do the same for C5.0.  Which one has better performance on the test data?
```{r}

library(C50)
tree5 <- C5.0(V1 ~ ., data=train, method="class")

```

```{r}
pred <- predict(tree5, test, type="class")
tab <- table(test$V1, pred)
tab
print(paste("Accuracy:", sum(diag(tab)) / sum(tab)))
```

Yes, it may get 100% accurate results on the test data, depending on how the test data was generated.

