---
title: "R Mapping Notebook"
output: html_notebook
name: "Mark Mitri"
date: 11 November 2022
---

### 0. Getting Started

Start by reviewing the slides and R notebooks from Wednesday's mapping lecture.  Our goal today is to build a choropleth map of PA counties showing median household incomes.  (Look at the end of the lab to see a sample target image.)

### 1. Find a county-level shapefile

Now go to the web and find county-level shapefiles that inlude PA counties.  Download the file (likely a zip file), and find the .shp file within it.  Read the shapefile using `st_read()`.
```{r}
library(ggplot2)
library(ggmap)
library(tidyverse)
library(sf)
library(mapview)
```

```{r}
marylandCounties <- st_read("C:\\Users\\MitriDell\\Documents\\Textbooks\\Senior - Fall 2022\\CSE 160\\Labs\\Final Project\\P5_3_2_counties_all_md.shp")
```

### 2. Limit it to PA counties

We only want to plot counties in PA.  So if the shapefile you have found is for more than PA, you'll need to extract only the PA data.
After extraction, how many counties do you have?  Does it match the correct number of counties for PA?

```{r}
#N/a
```

### 3. Plot PA counties

Use ggplot() and geom_sf() to plot the basic county outlines for PA.
```{r}
m0 <- ggplot(data=marylandCounties$geometry)
m1 <- m0 + geom_sf()
m1
```

```{r}
dataset <- read.csv("Traffic_Violations.csv", header = TRUE)
```

```{r}
plot(st_geometry(marylandCounties), xlim=c(0,42), ylim=c(-152,40))
points(x=mergedData$Latitude, y=mergedData$Longitude, col="red", cex=.5)
```

```{r}
mapview(marylandCounties, xcol = mergedData$Longitude, ycol = mergedData$Latitude)

```