---
title: "HW6"
author: "Your name"
date: "today"
output: html_notebook
---

(A total of 10 points.)

Consider the problem of time series analysis for parking data.  At the webpage https://archive.ics.uci.edu/ml/datasets/Parking+Birmingham you'll find a dataset about parking utilization.  You'll need to download the zip file and unzip it locally.

```{r}
parking <- read.csv("~/Dropbox/Courses/CSE160/data/parkingdata.csv")
parking$SystemCodeNumber <- factor(parking$SystemCodeNumber)
parking$LastDate <- as.Date(parking$LastUpdated)
parking$LastTime <- strptime(parking$LastUpdated, "%Y-%m-%d %H:%M:%S")
str(parking)
levels(parking$SystemCodeNumber)
```

After you have read in the data, and examined it, you'll find that it contains occupancy numbers for multiple locations.  Let's work with just one location: 
SystemCodeNumber == "BHMBCCMKT01" so generate a new dataframe by extracting the subset for this location.  (2 points)

Let's build a model of regression, to predict the next occupancy level reading.  Using the second time series analysis R notebook as reference, create a dataframe with multiple attributes (3 points).  Note that this dataset is quite different, and so not all of the attributes we saw in class could be created here.

```{r}
location1 <- parking[parking$SystemCodeNumber == "BHMBCCMKT01",]

parkingdf <- data.frame(Occupancy = location1$Occupancy, LastDate = location1$LastDate, LastTime = location1$LastTime)

parkingdf$Prior1[1] <- NA
for (i in 2:length(parkingdf$Occupancy)) {
  parkingdf$Prior1[i] <- parkingdf$Occupancy[i-1]
}

for (i in 1:18) {
  parkingdf$priorday[i] <- NA
}
for (i in 19:length(parkingdf$Occupancy)) {
  parkingdf$priorday[i] <- parkingdf$Occupancy[i-18]
}

parkingdf$id <- as.integer(rownames(parkingdf))

#rows <- dim(parkingdf)[1]
#rows

#train <- parkingdf[19:floor(rows*.8),]
#test <- parkingdf[-(1:floor(rows*.8)),]

#dim(train)
#dim(test)


```

Build a regression model based on those attributes (3 points) and evaluate it using the simple measures we saw in class (2 points).  Are you able to get 90% correlation?

```{r}
model <- glm(Occupancy ~ Prior1 + priorday + id, data=train)
summary(model)
pred.value <- predict(model, test)

# make actuals_predicteds dataframe.
actuals_preds <- data.frame(cbind(actuals=test$Occupancy, predicteds=pred.value))  
correlation_accuracy <- cor(actuals_preds$actuals, actuals_preds$predicteds)  # 63.7%
correlation_accuracy # want as high as possible

#head(actuals_preds)

mape <- mean(abs((actuals_preds$predicteds - actuals_preds$actuals))/actuals_preds$actuals)
mape # want as low as possible
mse <- mean((test$Occupancy-pred.value)^2)
mse # want as low as possible

# Yes we are able to get .90 correlation
```

Submission:

- Make sure this notebook is cleanly written so that someone can simply run the whole thing.  Submit this complete notebook via coursesite.
- Make sure to comment sufficiently that a reader will understand what you are doing.
